---
title: "FIT5145Assignment3"
author: "YuxiangZou"
date: "2024-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load library
```{r}
library(tidyverse)
library(readr)
library(stringr)
library(lubridate)
library(randomForest)
library(caret)
library(zoo)
library(xgboost)
library(dplyr)
```

#load csv
```{r}
weather_data <- read_csv("D:/FIT5145Assignment3/FIT5145Assignment3/melbourne_weather_data.csv")
agriculture_data <- read_csv("D:/FIT5145Assignment3/FIT5145Assignment3/agricultural_data.csv")
soil_data <- read_csv("D:/FIT5145Assignment3/FIT5145Assignment3/soil_data.csv")
```



#Descriptive statistics for agricultural data
```{r}
agriculture_summary <- agriculture_data %>% summarise(
  Wheat_Yield_Mean = mean(Wheat_yield, na.rm = TRUE),
  Wheat_Yield_SD = sd(Wheat_yield, na.rm = TRUE),
  Wheat_Yield_Min = min(Wheat_yield, na.rm = TRUE),
  Wheat_Yield_Max = max(Wheat_yield, na.rm = TRUE),
  Barley_Yield_Mean = mean(Barley_yield, na.rm = TRUE),
  Barley_Yield_SD = sd(Barley_yield, na.rm = TRUE),
  Barley_Yield_Min = min(Barley_yield, na.rm = TRUE),
  Barley_Yield_Max = max(Barley_yield, na.rm = TRUE),
  Rapeseed_Yield_Mean = mean(Rapeseed_yield, na.rm = TRUE),
  Rapeseed_Yield_SD = sd(Rapeseed_yield, na.rm = TRUE),
  Rapeseed_Yield_Min = min(Rapeseed_yield, na.rm = TRUE),
  Rapeseed_Yield_Max = max(Rapeseed_yield, na.rm = TRUE)
)
```
#Descriptive statistics for weather data
```{r}
weather_summary <- weather_data %>% summarise(
  Temperature_Mean = mean(Temperature, na.rm = TRUE),
  Temperature_SD = sd(Temperature, na.rm = TRUE),
  Temperature_Min = min(Temperature, na.rm = TRUE),
  Temperature_Max = max(Temperature, na.rm = TRUE),
  Precipitation_Mean = mean(Precipitation, na.rm = TRUE),
  Precipitation_SD = sd(Precipitation, na.rm = TRUE),
  Precipitation_Min = min(Precipitation, na.rm = TRUE),
  Precipitation_Max = max(Precipitation, na.rm = TRUE),
  Humidity_Mean = mean(Humidity, na.rm = TRUE),
  Humidity_SD = sd(Humidity, na.rm = TRUE),
  Humidity_Min = min(Humidity, na.rm = TRUE),
  Humidity_Max = max(Humidity, na.rm = TRUE)
)
```


#Descriptive Statistics for Soil Data
```{r}
soil_summary <- soil_data %>% summarise(
  Soil_pH_Mean = mean(Soil_pH, na.rm = TRUE),
  Soil_pH_SD = sd(Soil_pH, na.rm = TRUE),
  Soil_pH_Min = min(Soil_pH, na.rm = TRUE),
  Soil_pH_Max = max(Soil_pH, na.rm = TRUE),
  Organic_Matter_Mean = mean(Organic_matter, na.rm = TRUE),
  Organic_Matter_SD = sd(Organic_matter, na.rm = TRUE),
  Organic_Matter_Min = min(Organic_matter, na.rm = TRUE),
  Organic_Matter_Max = max(Organic_matter, na.rm = TRUE),
  Nitrogen_Mean = mean(Nitrogen, na.rm = TRUE),
  Nitrogen_SD = sd(Nitrogen, na.rm = TRUE),
  Nitrogen_Min = min(Nitrogen, na.rm = TRUE),
  Nitrogen_Max = max(Nitrogen, na.rm = TRUE),
  Phosphorus_Mean = mean(Phosphorus, na.rm = TRUE),
  Phosphorus_SD = sd(Phosphorus, na.rm = TRUE),
  Phosphorus_Min = min(Phosphorus, na.rm = TRUE),
  Phosphorus_Max = max(Phosphorus, na.rm = TRUE),
  Potassium_Mean = mean(Potassium, na.rm = TRUE),
  Potassium_SD = sd(Potassium, na.rm = TRUE),
  Potassium_Min = min(Potassium, na.rm = TRUE),
  Potassium_Max = max(Potassium, na.rm = TRUE)
)
```

#Bar chart of wheat production over the years
```{r}
library(ggplot2)

ggplot(agriculture_data, aes(x = as.Date(Date), y = Wheat_yield, fill = Wheat_yield)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "Wheat Yield Over Years", x = "Year", y = "Wheat Yield", fill = "Yield") +
  theme_minimal()


```
#Rapeseed_Yield line chart
```{r}
ggplot(agriculture_data, aes(x = as.Date(Date), y = Rapeseed_yield)) +
  geom_line(color = "darkgreen") +
  labs(title = "Rapeseed Yield Over Years", x = "Year", y = "Rapeseed Yield") +
  theme_minimal()
```

#Barley_Yield line chart
```{r}
ggplot(agriculture_data, aes(x = as.Date(Date), y = Barley_yield)) +
  geom_line(color = "darkblue") +
  labs(title = "Barley Yield Over Years", x = "Year", y = "Barley Yield") +
  theme_minimal()
```









#Weather data line chart
```{r}
ggplot(weather_data, aes(x = as.Date(Date))) +
  geom_line(aes(y = Temperature, color = "Temperature"), linewidth = 1) +
  geom_line(aes(y = Precipitation, color = "Precipitation"), linewidth = 1) +
  geom_line(aes(y = Humidity, color = "Humidity"), linewidth = 1) +
  scale_color_manual(values = c("Temperature" = "blue",
                                "Precipitation" = "green",
                                "Humidity" = "red")) +
  labs(title = "Weather Data Trends",
       x = "Date",
       y = "Value",
       color = "Parameters") +
  theme_minimal()
```
#Line chart of soil data
```{r}
ggplot(soil_data, aes(x = as.Date(Date))) +
  geom_line(aes(y = Soil_pH, color = "Soil pH"), linewidth = 1) +
  geom_line(aes(y = Organic_matter, color = "Organic Matter"), linewidth = 1) +
  geom_line(aes(y = Nitrogen, color = "Nitrogen"), linewidth = 1) +
  geom_line(aes(y = Phosphorus, color = "Phosphorus"), linewidth = 1) +
  geom_line(aes(y = Potassium, color = "Potassium"), linewidth = 1) +
  scale_color_manual(values = c("Soil pH" = "blue",
                                "Organic Matter" = "green",
                                "Nitrogen" = "red",
                                "Phosphorus" = "purple",
                                "Potassium" = "orange")) +
  labs(title = "Soil Data Trends",
       x = "Date",
       y = "Value",
       color = "Parameters") +
  theme_minimal()
```

#Using random forest regression analysis


#convert date format
```{r}
agriculture_data$Date <- as.Date(agriculture_data$Date)
weather_data$Date <- as.Date(weather_data$Date)
soil_data$Date <- as.Date(soil_data$Date)
```

#Extract year and month information
```{r}
weather_data <- weather_data %>%
  mutate(Year = year(Date), Month = month(Date))
soil_data <- soil_data %>%
  mutate(Year = year(Date), Month = month(Date))
```

#Aggregate weather data to monthly 
```{r}
monthly_weather_data <- weather_data %>%
  group_by(Year, Month) %>%
  summarize(Temperature = mean(Temperature, na.rm = TRUE),
            Precipitation = mean(Precipitation, na.rm = TRUE),
            Humidity = mean(Humidity, na.rm = TRUE))
```

#monthly soil data
```{r}
monthly_soil_data <- soil_data %>%
  group_by(Year, Month, Region) %>%
  summarize(Soil_pH = mean(Soil_pH, na.rm = TRUE),
            Organic_matter = mean(Organic_matter, na.rm = TRUE),
            Nitrogen = mean(Nitrogen, na.rm = TRUE),
            Phosphorus = mean(Phosphorus, na.rm = TRUE),
            Potassium = mean(Potassium, na.rm = TRUE))
```

#Assume monthly weightings and downgrade to monthly production based on annual production trends
```{r}
wheat_weights <- c(0.1, 0.02, 0.03, 0.03, 0.02, 0.02, 0.03, 0.02, 0.03, 0.08, 0.35, 0.25)
barley_weights <- c(0.07, 0.03, 0.1, 0.07, 0.05, 0.04, 0.11, 0.09, 0.08, 0.07, 0.15, 0.14)
rapeseed_weights <- c(0.06, 0.04, 0.07, 0.1, 0.04, 0.05, 0.09, 0.11, 0.07, 0.08, 0.13, 0.16)
monthly_agriculture_data <- agriculture_data %>%
  mutate(Year = year(Date)) %>%
  uncount(weights = 12, .remove = FALSE) %>%
  mutate(Month = rep(1:12, times = nrow(agriculture_data))) %>%
  group_by(Year) %>%
  mutate(
    Wheat_yield = Wheat_yield * wheat_weights[Month],
    Barley_yield = Barley_yield * barley_weights[Month],  
    Rapeseed_yield = Rapeseed_yield * rapeseed_weights[Month] 
  ) %>%
  ungroup() %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-")))

```

#merge monthly soil and weather data
```{r}
monthly_data <- monthly_agriculture_data %>%
  left_join(monthly_weather_data, by = c("Year", "Month")) %>%
  left_join(monthly_soil_data, by = c("Year", "Month"))
monthly_data <- na.omit(monthly_data)
```

## Split the data set into training set and test set
```{r}
set.seed(123)
train_indices <- createDataPartition(monthly_data$Wheat_yield, p = 0.8, list = FALSE)
train_data <- monthly_data[train_indices, ]
test_data <- monthly_data[-train_indices, ]
```

#Train random forest regression model
```{r}
rf_model <- randomForest(Wheat_yield ~ Temperature + Precipitation + Humidity + Soil_pH + Organic_matter + Nitrogen + Phosphorus + Potassium, data = train_data, ntree = 500, mtry = 3)
```


#Predicting one month future agricultural yields
```{r}
future_weather <- data.frame(
  Month = as.Date("2025-01-01"), 
  Temperature = 17,  
  Precipitation =4 ,  
  Humidity = 85  
)

future_soil <- data.frame(
  Month = as.Date("2025-01-01"),  
  Soil_pH = 5.3,  
  Organic_matter = 2.7, 
  Nitrogen = 0.88,  
  Phosphorus = 0.08, 
  Potassium = 0.20  
)

future_data <- future_soil %>%
  left_join(future_weather, by = "Month")

future_prediction <- predict(rf_model, newdata = future_data)
cat("Predicted Wheat Yield: ", future_prediction, "\n")
```
#predict one year
```{r}


# Set weather data for the next year
future_weather_year <- data.frame(
  Month = seq(as.Date("2025-01-01"), as.Date("2025-12-01"), by = "month"),
  Temperature = c(17, 16, 18, 17, 19, 17, 17, 16, 17, 16, 17, 15),  
  Precipitation = c(4, 9, 10, 8, 7, 10, 11, 9, 7, 8, 8, 8),            
  Humidity = c(85, 69, 73, 66, 64, 62, 61, 66, 67, 68, 65, 64)      
)

#set soil data for the next year
future_soil_year <- data.frame(
  Month = seq(as.Date("2025-01-01"), as.Date("2025-12-01"), by = "month"),
  Soil_pH = c(5.2, 8.3, 7.4, 6.8, 5.12, 5.13, 4.73, 7.96, 6.9, 7.3, 5.4, 8.3),  
  Organic_matter = c(2.7, 2.7, 4.7, 1.9, 2.8, 3.1, 3.2, 4.1, 1.3, 2.7, 1.8, 1.7),  
  Nitrogen = c(0.85, 0.86, 0.87, 0.88, 0.43, 0.36, 0.82, 0.56, 0.55, 0.81, 0.87, 0.86),  
  Phosphorus = c(0.07, 0.08, 0.09, 0.07, 0.11, 0.12, 0.13, 0.12, 0.11, 0.10, 0.09, 0.08),  
  Potassium = c(0.20, 1.72, 1.20, 1.21, 1.22, 0.83, 0.54, 0.95, 0.34, 1.91, 1.37, 1.25)  
)

#Merging future weather and soil data
future_year_data <- future_soil_year %>%
  left_join(future_weather_year, by = "Month")

#Forecasting agricultural production for the next year
future_year_predictions <- predict(rf_model, newdata = future_year_data)

#Add prediction results to the data frame
future_year_data$Predicted_Yield <- future_year_predictions

# chart
ggplot(future_year_data, aes(x = Month, y = Predicted_Yield)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Predicted wheat Yield for 2025", x = "Month", y = "Predicted Yield") +
  theme_minimal()

```


